---
title: "Use cases - Thermal limits"
output: github_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(arrow)
library(terra)
knitr::opts_chunk$set(echo = FALSE)
species_names <- c(
  "Cephalopholis argus",
  "Dascyllus aruanus",
  "Thalassoma lucasanum",
  "Echinometra lucunter",
  "Uca maracoani"
)

# Load OBIS full export to see number of records for each species
obis_ds <- open_dataset("~/Research/mpa_europe/mpaeu_shared/obis_20231025.parquet")

obis_filt <- obis_ds %>%
  filter(species %in% species_names) %>%
  select(AphiaID, species,
         occurrenceID, datasetID,
         minimumDepthInMeters, maximumDepthInMeters,
         decimalLongitude, decimalLatitude,
         eventDate) %>%
  collect()

obis_filt$unID <- 1:nrow(obis_filt)

# Load bio-oracle
bio_or <- rast("~/Research/mpa_europe/mpaeu_sdm/data/env/current/thetao_baseline_depthsurf_mean.tif")

sst_vals <- terra::extract(bio_or, obis_filt[,c("decimalLongitude", "decimalLatitude")], ID = F)

which_na <- which(is.na(sst_vals[,1]))
cell_na <- cellFromXY(bio_or, as.data.frame(obis_filt[which_na, c("decimalLongitude", "decimalLatitude")]))
adj_cells <- adjacent(bio_or, cell_na, directions = "queen")
adj_vals <- unname(unlist(apply(adj_cells, 1, function(x){
  zv <- mean(terra::extract(bio_or, x)[,1], na.rm = T)
  if (is.null(zv) | length(zv) < 1) {
    NA
  } else {
    zv
  }
})))
sst_vals[which_na,1] <- adj_vals

obis_filt <- cbind(obis_filt, sst_biooracle = sst_vals[,1])
```

## Fishes

### _Cephalopholis argus_

Estimated thermal optima: 28    
Estimated thermal upper limit: 29.5    
Experimental thermal optima: NA   
Experimental thermal upper limit: NA    

```{r}
sp <- "Cephalopholis argus"
obis_filt %>%
  filter(species == sp) %>%
  filter(!is.na(sst_biooracle)) %>%
  ggplot() +
  geom_density(aes(x = sst_biooracle), fill = "#BCCBD2", color = "#266E92") +
  ggtitle(sp)+xlab("SST (°C)") +
  theme_light()
```

### _Dascyllus aruanus_

Estimated thermal optima: 25.6     
Estimated thermal upper limit: 29      
Experimental thermal optima: NA    
Experimental thermal upper limit: 40.5      

```{r}
sp <- "Dascyllus aruanus"
obis_filt %>%
  filter(species == sp) %>%
  filter(!is.na(sst_biooracle)) %>%
  ggplot() +
  geom_density(aes(x = sst_biooracle), fill = "#BCCBD2", color = "#266E92") +
  ggtitle(sp)+xlab("SST (°C)") +
  theme_light()
```

### _Thalassoma lucasanum_

Estimated thermal optima: 28.8       
Estimated thermal upper limit: 29.2      
Experimental thermal optima: NA      
Experimental thermal upper limit: 36      

```{r}
sp <- "Thalassoma lucasanum"
obis_filt %>%
  filter(species == sp) %>%
  filter(!is.na(sst_biooracle)) %>%
  ggplot() +
  geom_density(aes(x = sst_biooracle), fill = "#BCCBD2", color = "#266E92") +
  ggtitle(sp)+xlab("SST (°C)") +
  theme_light()
```



## Invertebrates

### _Echinometra lucunter_

Estimated thermal optima: NA    
Estimated thermal upper limit: NA    
Experimental thermal optima: 29.4    
Experimental thermal upper limit: 36    


```{r}
sp <- "Echinometra lucunter"
obis_filt %>%
  filter(species == sp) %>%
  filter(!is.na(sst_biooracle)) %>%
  ggplot() +
  geom_density(aes(x = sst_biooracle), fill = "#BCCBD2", color = "#266E92") +
  ggtitle(sp)+xlab("SST (°C)") +
  theme_light()
```

### _Uca maracoani_

Estimated thermal optima: NA    
Estimated thermal upper limit: NA    
Experimental thermal optima: NA    
Experimental thermal upper limit: 38.6     

```{r}
sp <- "Uca maracoani"
obis_filt %>%
  filter(species == sp) %>%
  filter(!is.na(sst_biooracle)) %>%
  ggplot() +
  geom_density(aes(x = sst_biooracle), fill = "#BCCBD2", color = "#266E92") +
  ggtitle(sp)+xlab("SST (°C)") +
  theme_light()
```

